Homicide Increases in the COVID-19 Era

### 2025-03-04 Draft2.qmd
hyperlinks

The codebook for this dataset can be found here: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/about_data.

@@

## Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025
```{r, echo=FALSE}
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

@@

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_stata(scheme = "s1mono") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```
### Interpretation
This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data. By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions. Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month. There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average. Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.

@@

ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
        geom_boxplot(aes(fill = Bureau)) +
  labs(y = "Annual Number of Crimes")
```
### Interpretation

@@

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) +
  labs(fill = "Correlation")
```
### Interpretation
This heat map shows the correlations between different variables that could affect the number of crimes committed. Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles. The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.

@@

levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
  geom_jitter(alpha = 0.4, width = 0.2) +
  labs(x = "Gender",
       y = "Age") +
  theme_minimal() +
  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
  theme(legend.position = 'none')
```
### Interpretation

@@

ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(x = "Hour (Military Time)",
       y = "Density") +
  theme_minimal() +
  theme(legend.position = 'none')
```
### Interpretation

@@

### Interpretation
Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic. In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic. On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.

@@

# Data Sources {.unnumbered}
## Primary Datasets {.unnumbered}
Crime Data From 2010-2019: https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data

Crime Data From 2020-Present: https://catalog.data.gov/dataset/crime-data-from-2020-to-present

## Secondary Datasets {.unnumbered}
Los Angeles County Eviction Data: https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/

City of Los Angeles Eviction Data: https://housing.lacity.gov/residents/renters/eviction-notices-filed

Median Income Data: https://data.census.gov/table/ACSST1Y2023.S1903?q=income

@@

##### Zip Codes for Crimes
```{r Get Zip Codes, eval=FALSE}
library(sf)
library(dplyr)

@@

# creating labels
levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
  labs(
    x = "Year",
    y = "Number of Homicides",
    fill = "Income Quartile",
    title = "Homicide Rates in Los Angeles By Median Income Level"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma", direction = -1)
```
#### Crime \~ Week Regression Code {.unnumbered}

>>>>

The codebook for this dataset can be found here: <https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/about_data>.

@@

## Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025

```{r, echo=FALSE}
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

@@

This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data. By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions. Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month. There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average. Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.

@@

@@

This heat map shows the correlations between different variables that could affect the number of crimes committed. Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles. The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.

@@

@@

@@

Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic. In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic. On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.

@@

# Data Sources {.unnumbered}

## Primary Datasets {.unnumbered}

Crime Data From 2010-2019: <https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data>

Crime Data From 2020-Present: <https://catalog.data.gov/dataset/crime-data-from-2020-to-present>

## Secondary Datasets {.unnumbered}

Los Angeles County Eviction Data: <https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/>

City of Los Angeles Eviction Data: <https://housing.lacity.gov/residents/renters/eviction-notices-filed>

Median Income Data: <https://data.census.gov/table/ACSST1Y2023.S1903?q=income>

@@

##### Zip Codes for Crimes

```{r Get Zip Codes, eval=FALSE}
library(sf)
library(dplyr)

@@

#### Crime \~ Week Regression Code {.unnumbered}
